---
title: "Cognitive Modeling LC1"
author: "Harm Manders, 10677186"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

***
```{R}
## set working direcotry
setwd("/home/harm/Uni/cogmod/LC1/") ##<======= SET THIS TO YOUR LC1 FOLDER!
```

***  

## ANSWER SECTION 


***

### Q1 Try _?log_ to read how to use log function. What are the two ways to calculate the log of 10 using base 10?
```{r, eval=FALSE}
log(10, 10)
log10(10)
```
***  
  
### Q2 if you assing an integer value to a variable ( e.g. k = 1), will the data value also be integer  

No, k will be of class numeric
```{r}
k = 1
class(k)
```

***  

### Q3 the __seq__ function is very useful ot make sequences of numbers, which will be saved as vectors. Look up how __seq__ works and make a sequence starting at 0 ending at 100 in steps of 2?  
```{r}
seq(0,100,2)
```

*** 
### Q4 find the minimum and maximum of the random numbers you just generated, and also the position in the list:
```{r}
set.seed(123)
rr =runif(100, 0, 1) 
max(rr)
which.max(rr)
min(rr)
which.min(rr)
```
 
***
### Q5 try to plot all random numbers using __plot()__ and identify the min and max in it using __abline()__ to add (h)orizontal and (v)ertical lines 
(hint: use ?plot and ?abline)  
```{r}
plot(rr)
abline(h = max(rr), v = which.max(rr))
abline(h = min(rr), v = which.min(rr))
```

***
### Q6 Below are two lists, can you combine them using __rbind__, if not, why not and how could you change that?
No, for _rbind_ the number of columns must be the same. Thus change ncol to same value
```{r}
L1 = matrix( c(2, 4, 3, 1, 5, 7), nrow=3, ncol=2, byrow = TRUE)        
L2  = matrix( c(2, 4, 3, 1, 5, 7), nrow=2, ncol=2, byrow = F)
rbind(L1,L2)

```

***
### Q7  can you now select all cars that drove more that 30 kilometers in distance? 
```{r}
test = subset(cars, cars$dist > 30)
# First 10 rows
test[1:10,]
```

***
### Q8  can you use ifelse to tell for each element in the following list if it is even or odd ? (hint modulo: x %% y)
```{r, eval=T}
a = c(5,7,2,9)
ifelse(a %% 2 == 0, 1, 0)
```